Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2012-10-02 Dail_Mad_ext.tex
discussion

  In some respects, the DM model can be viewed as a specific
  manifestation of the general state-space model formulated by
  \citep{buckland_etal:2004}. However, their model was designed to
  allow for inference about demographic parameters, which will
  typically require mark-recapture data. Although this may be viewed
  as the gold standard in many studies of wildlife populations, in
  this paper, we focus on the analysis of count data, which is the
  typical source of information used in state-state modeling
  applications, and one of the most widely available forms of
  ecological data---it is also the data required by the DM model.
\end{comment}

Our work has highlighted the limitations of classical state-space
models as applied to ecological time series data, and we have
demonstrated how open population $N$-mixture models overcome these
limitations. We extended this class of models in several important
ways, while keeping in mind the practical issues ...

@@

@@

The zero-inflated models for initial
abundance and dynamics allow one to estimate and model the proportion
of sites outside of the range of a species.  This could be especially
useful when different factors control the range than the abundance or
dynamics within the range.  [some examples where that is the case?]

[paragraph on random effects]

@@

\begin{comment}
Should we include these possible extensions as well?

@@

>>>>

\end{comment}

Our work has highlighted the limitations of classical state-space
models as applied to ecological time series data, and we have
demonstrated how open population $N$-mixture models overcome these
limitations. We extended this class of models in several important
ways that should make them more suitable to typical ecological
datasets that are characterized by sparse counts, zero-inflation, and
observation error due to imperfect detection.

@@

In the context of IPMs, the DM model can be viewed as a specific
manifestation of the general state-space model formulated by
\citep{buckland_etal:2004}. However, their model was designed to
allow for inference about demographic parameters, which will
typically require mark-recapture data. Although this may be viewed
as the gold standard in many studies of wildlife populations, in
this paper, we focus on the analysis of count data, which is the
typical source of information used in state-state modeling
applications, and one of the most widely available forms of
ecological data---it is also the data required by the DM model.

@@

\begin{comment}
  The zero-inflated models for initial abundance and dynamics allow
  one to estimate and model the proportion of sites outside of the
  range of a species.  This could be especially useful when different
  factors control the range than the abundance or dynamics within the
  range.  [some examples where that is the case?]
\end{comment}
[paragraph on random effects]

@@

\begin{comment}

@@

\section{Acknowledgements}

We thank J. Andrew Royle and T. Scott Sillett for helpful comments
throughout this process.
